---
title: "Iconicity Analyses"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

###Load required packages
```{r results='hide', warning=FALSE, message=FALSE}
packages = c("tidyverse","ggplot2","beeswarm","colorspace", "stargazer", "MASS", "ordinal", "generalhoslem", "qwraps2", "gridExtra","ggpubr","sure", "VGAM", "truncreg", "car", "rmarkdown", "table1", "kableExtra", "epitools")
package.check <- lapply(packages, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

#verify they are loaded
search()

```

###Check and then set working dir
```{r results='hide', echo=FALSE}
setwd("/Users/emilycarrigan/Dropbox/Data Analysis Work") ##this should be wherever your file is saved 

GiveN <- read.csv("GN_Smsummary_Lgall_200409.csv", na.strings = c("N/A", "I/R", "", "TBD")) #import the data file (which I already saved as a csv from excel file)
typeof(GiveN) # when importing using read.csv, resulting obj type is a list
View(GiveN)
```

###DATA CLEANING/PREP/CHECK
```{r}
GiveN_subset <- GiveN[1:237,] #if you need to remove blank rows/unnecessary columns in the dataframe, subset by [row,column] ranges
View(GiveN_subset)
GiveN_subset <- subset(GiveN_subset, Including.in.Study.=='Yes' & Including_GiveN.=='Yes' & GN_Small_Ceiling._FinalSet_conservative!='') # This can be amended as necessary for the specific analyses 
View(GiveN_subset)

str(GiveN_subset) # check the data
#Age and SES SHOULD import as 'num' type (bc decimal points)
# All GiveN values should import as integer type (but we will convert them to factor below)
## If they import as something other than integer, that typically means there is text in one of the cells in that column OTHER than the "N/A" string
summary(GiveN_subset) # check whether have missing values for any variables

#reorder factor levels for Language Group so they show up in order I want them
GiveN_subset$LanguageGroup <- as.factor(factor(as.character(GiveN_subset$Group_4cat), levels = c("English Early", "ASL Early", "English Later", "ASL Later"), exclude=NA))

#BELOW uses names in MASTER spreadsheet
GiveN_subset$GN_sm_fac_c <- as.factor(GiveN_subset$GN_Small_Ceiling._FinalSet_conservative)
GiveN_subset$GN_ALL_sm_fac <- as.factor(GiveN_subset$GN_Small_Ceiling._FinalSet)


#Create Modality and Timing variables and add to dataframe
GiveN_subset$Language_Modality <- ifelse(GiveN_subset$Group_4cat == "English Early" | GiveN_subset$Group_4cat == "English Later", "English", "ASL")

GiveN_subset$Language_Timing <- factor(as.character(GiveN_subset$Group_2cat), levels = c("Early", "Later"), exclude="")
##ORDER MATTERS - I initially acidentally ordered things "Later", "Early", which actually reversed the automatically alphabetically ordered variables in the original Group2Cat (and labeled the opposite of what they actually were--didn't realize until I ran the models and was confused about why it seemed like Early exposed kids were doing worse--YIKES!)

#subset only Early-exposed participants
GiveN_Early <- subset(GiveN_subset, GiveN_subset$Language_Timing=="Early" & GiveN_subset$Age>=3.4)
#removing the kids in English early under Age 3.4 (youngest age for ASL early) to age match the samples a bit better only removes 3 kids (now 101 instead of 104)
View(GiveN_Early)
```

###ASSIGN shorter VARIABLE names
```{r}
#outcome vars (these two are the same variable, but represented as either an integer or a factor)
GN_sm_int <- GiveN_Early$GN_Small_Ceiling._FinalSet_conservative #KL drawing only from Give-N small data (which all children completed)
GN_sm_fac <- GiveN_Early$GN_sm_fac_c

#predictors
Age <- GiveN_Early$Age
SES <- GiveN_Early$SES..8.66.
LangGrp <- GiveN_Early$LanguageGroup
Count_NObj <- GiveN_Early$Highest_Count_noobj
Count_WObj <- GiveN_Early$Highest_Count_wobj
MAX_count_seq <- GiveN_Early$MAX_COUNT_SEQ
Modality <- GiveN_Early$Language_Modality
Timing <- GiveN_Early$Language_Timing

```


###Demographics & Basic plots
```{r}
summary(Age)
sd(Age)
summary(Age >7) #tells you how many kids over age 7
summary(Age ==7)
summary(Age=='')#check for missing values
summary(SES)
sd(SES)
summary(SES=='')#check for missing values

summary(LangGrp)
summary(Timing)
summary(Modality)
summary(Count_NObj)

#TO GET AGES/SES etc by group
aggregate(formula = GiveN_Early$Age~GiveN_Early$LanguageGroup, FUN = mean) # gives means by group!!!
aggregate(formula = GiveN_Early$Age~GiveN_Early$LanguageGroup, FUN = min) # gives means by group!!!
aggregate(formula = GiveN_Early$Age~GiveN_Early$LanguageGroup, FUN = max) # gives means by group!!!
aggregate(formula = GiveN_Early$Age~GiveN_Early$LanguageGroup, FUN = sd) # gives means by group!!!

ggplot(data=subset(GiveN_Early, !is.na(LangGrp)), aes(x=LangGrp, y=Age)) + geom_violin() + labs(x="Language Group", y="Age (years)")


aggregate(formula = GiveN_Early$SES..8.66. ~ GiveN_Early$LanguageGroup, FUN = mean) # gives means by group!!!
aggregate(formula = GiveN_Early$SES..8.66. ~ GiveN_Early$LanguageGroup, FUN = min) # gives means by group!!!
aggregate(formula = GiveN_Early$SES..8.66. ~ GiveN_Early$LanguageGroup, FUN = max) # gives means by group!!!
aggregate(formula = GiveN_Early$SES..8.66. ~ GiveN_Early$LanguageGroup, FUN = sd) # gives means by group!!!

jitter <- position_jitter(width=0.2, height=0.1)

ggplot(data=subset(GiveN_Early, (Language_Modality=="ASL")), aes(x=SES..8.66., y=..count..)) + geom_density()
ggplot(data=subset(GiveN_Early, (Language_Modality=="English")), aes(x=SES..8.66., y=..count..)) + geom_density()

ggplot(data=subset(GiveN_Early, (Language_Modality=="ASL")), aes(x=Age, y=..count..)) + geom_density()
ggplot(data=subset(GiveN_Early, (Language_Modality=="English")), aes(x=Age, y=..count..)) + geom_density()

ggplot(data=subset(GiveN_Early, !is.na(LangGrp)), aes(x=Age, y=GN_sm_int)) + geom_point(aes(shape=Modality, color=Modality), position=jitter, size=2.5) + geom_smooth(mapping=aes(group=Modality, color=Modality), method="loess", se=FALSE) + labs(x="Age (years)", y="Give-N highest number correctly given (up to 6)")
ggplot(data=subset(GiveN_Early, !is.na(LangGrp)), aes(x=Age, y=GN_sm_int)) + geom_jitter(aes(shape=Modality, color=Modality), size=2.5) + geom_smooth(mapping=aes(group=Modality, color=Modality), method="loess", se=FALSE) + labs(x="Age (years)", y="Give-N highest number correctly given (up to 6)")
```

###Subsetting the data by age
We have quite a few participants at ceiling on Give-N, likely because of our age range, and this may be obscuring any potential group differences--so we subset to a smaller age range, using information on the distribution of ages
```{r}
##TABLE OF AGES
#set up age bins first
# set up cut-off values 
breaks <- c(3.4, 4, 4.5, 5, 5.5, 6, 6.5, 7, 7.6)
# specify interval/bin labels
tags <- c("[3.4-4)","[4-4.5)", "[4.5-5)", "[5-5.5)", "[5.5-6)", "[6-6.5)","[6.5-7)", "[7-7.6)")
# bucketing values into bins
GiveN_Early$AgeBins <- cut(GiveN_Early$Age, 
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)
# inspect bins
summary(GiveN_Early$AgeBins)

##make the tables
table(GiveN_Early$AgeBins, GiveN_Early$Language_Modality)
prop.table(table(GiveN_Early$AgeBins, GiveN_Early$Language_Modality))

#set up DIFFERENT age bins first
# set up cut-off values 
breaks_fewer<- c(3.4, 4.5, 5.5, 6.5, 7.6)
# specify interval/bin labels
tags_fewer <- c("[3.4-4.5)","[4.5-5.5)", "[5.5-6.5)", "[6.5-7.6)")
# bucketing values into bins
GiveN_Early$AgeBinsFewer <- cut(GiveN_Early$Age, 
                           breaks=breaks_fewer, 
                           include.lowest=TRUE, 
                           right=FALSE, 
                           labels=tags_fewer)
# inspect bins
summary(GiveN_Early$AgeBinsFewer)

##make the tables
table(GiveN_Early$AgeBinsFewer, GiveN_Early$Language_Modality)
prop.table(table(GiveN_Early$AgeBinsFewer, GiveN_Early$Language_Modality))
```

Based on charts and participant, we decided to subset the data to participants under 5.5 years of age
```{r}
GiveN_3_to_5_half <- subset(GiveN_Early, Age <5.6) 
View(GiveN_3_to_5_half)

##Contingency table for do kids know 5 or not by group
GiveN_3_to_5_half <- GiveN_3_to_5_half %>% dplyr::mutate(At_5 = ifelse(GN_Small_Ceiling._FinalSet_conservative>=5, "Yes", "No"))
View(GiveN_3_to_5_half)

##Table of how many kids "are "know 5" (or higher) by group (Contingency table)
table1::label(GiveN_3_to_5_half$At_5) <- "Does child know up to 5?"
table1::table1(~At_5 | Language_Modality, data = GiveN_3_to_5_half)

```

##QUESTION: Within this age group, do we see different proportions of kids who are 5-knowers vs. not

```{r}

#Using risk ratio code to see if English group sig different from ASL group
Early_ASL_v_English_know5 <- matrix(c(8,12,19,36), 2, 2, byrow=TRUE)
dimnames(Early_ASL_v_English_know5) <- list("Group" = c("Early ASL", "Early English (Typically Hearing)"), "Does child know up to 5?" = c("No", "Yes"))
Early_ASL_v_English_know5
riskratio.small(Early_ASL_v_English_know5)

```


Risk ratio analysis indicates that Early English participants aged 3.5-5.5 years are equally likely to be 5-knowers as Early ASL participants