---
title: The Trajectory of Early Lexical Acquisition is Invariant to Iconicity and Modality: The Case of Number Words
author:
- Clifton, Langdon, Ph.D.
- Emily Carrigan, Ph.D. (code author)
- Marie Coppola, Ph.D
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
date: "Code last edited  `r format(Sys.time(), '%d %m, %Y')`"
---





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

### Load required packages
```{r results='hide', warning=FALSE, message=FALSE}
#from https://www.vikram-baliga.com/blog/2015/7/19/a-hassle-free-way-to-verify-that-r-packages-are-installed-and-loaded

packages = c("tidyverse","ggplot2","beeswarm","colorspace", "stargazer", "MASS", "ordinal", "generalhoslem", "qwraps2", "gridExtra","ggpubr","sure", "VGAM", "truncreg", "car", "rmarkdown", "table1", "kableExtra", "epitools")
package.check <- lapply(packages, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})
```

### Check and then set working dir
```{r results="hide", echo=FALSE}
# setwd("") ##this should be wherever your file is saved 

GiveN <- read.csv("GN_Smsummary_Lgall_KW_201106.csv", na.strings = c("N/A", "I/R", "", "TBD", "N/R")) #import the data file (which I already saved as a csv from excel file)
typeof(GiveN) # when importing using read.csv, resulting obj type is a list
View(GiveN)
```

### DATA CLEANING/PREP/CHECK
```{r results="hide", echo=FALSE, include=FALSE}
GiveN_subset <- GiveN[1:237,] #if you need to remove blank rows/unnecessary columns in the dataframe, subset by [row,column] ranges
View(GiveN_subset)
GiveN_subset <- subset(GiveN_subset, Including.in.Study.=='Yes' & Including_GiveN.=='Yes' & GN_Small_Ceiling._FinalSet_conservative!='') # This can be amended as necessary for the specific analyses 
View(GiveN_subset)

str(GiveN_subset) # check the data
#Age and SES SHOULD import as 'num' type (bc decimal points)
# All GiveN values should import as integer type (but we will convert them to factor below)
## If they import as something other than integer, that typically means there is text in one of the cells in that column OTHER than the "N/A" string
summary(GiveN_subset) # check whether have missing values for any variables

#reorder factor levels for Language Group so they show up in order I want them
GiveN_subset$LanguageGroup <- as.factor(factor(as.character(GiveN_subset$Group_4cat), levels = c("English Early", "ASL Early", "English Later", "ASL Later"), exclude=NA))


#Create Modality and Timing variables and add to dataframe
GiveN_subset$Language_Modality <- ifelse(GiveN_subset$Group_4cat == "English Early" | GiveN_subset$Group_4cat == "English Later", "English", "ASL")

GiveN_subset$Language_Timing <- factor(as.character(GiveN_subset$Group_2cat), levels = c("Early", "Later"), exclude="")

GiveN_subset <-  dplyr::mutate(GiveN_subset, highestcorr_to5 = ifelse(GN_Small_Ceiling._FinalSet_conservative==6, 5, GN_Small_Ceiling._FinalSet_conservative))

GiveN_subset$GN_smto5_fac <- as.factor(GiveN_subset$highestcorr_to5)

##CREATE PERCENT correct Give-N LARGE measure (second analysis) - drawn from data in 5 separate columns 
GiveN_subset <-  dplyr::mutate(GiveN_subset, t7_corr = ifelse(is.na(GiveN_subset$Target7_FinalSet_Correct.), GiveN_subset$Target7_1st_Response_Fish_Correct., GiveN_subset$Target7_FinalSet_Correct.))
GiveN_subset <-  dplyr::mutate(GiveN_subset, t9_corr = ifelse(is.na(GiveN_subset$Target9_FinalSet_Correct.), GiveN_subset$Target9_1st_Response_Fish_Correct., GiveN_subset$Target9_FinalSet_Correct.))
GiveN_subset <-  dplyr::mutate(GiveN_subset, t10_corr = ifelse(is.na(GiveN_subset$Target10_FinalSet_Correct.), GiveN_subset$Target10_1st_Response_Fish_Correct., GiveN_subset$Target10_FinalSet_Correct.))
GiveN_subset <-  dplyr::mutate(GiveN_subset, t12_corr = ifelse(is.na(GiveN_subset$Target12_FinalSet_Correct.), GiveN_subset$Target12_1st_Response_Fish_Correct., GiveN_subset$Target12_FinalSet_Correct.))
GiveN_subset <-  dplyr::mutate(GiveN_subset, t16_corr = ifelse(is.na(GiveN_subset$Target16_FinalSet_Correct.), GiveN_subset$Target16_1st_Response_Fish_Correct., GiveN_subset$Target16_FinalSet_Correct.))


#Calculate mean proportion correct for Give-N trials 7, 9, 10, 12, & 16 (only for kids who got all five trials)

# Code below will appropriately *not* calculate mean (puts "NA") if any trial from above columns has NA value--THIS IS CORRECT, bc doing a proportion correct out of different numbers of trials for different kids is a no-go
GiveN_subset <-  GiveN_subset %>% rowwise() %>% dplyr::mutate(GNLg_propcorr_NA=mean(c(t7_corr, t9_corr, t10_corr, t12_corr, t16_corr)))
```

### Subset to Early-exposed participants only
```{r}
#subset only Early-exposed participants
GiveN_Early <- subset(GiveN_subset, GiveN_subset$Language_Timing=="Early" & GiveN_subset$Age>=3.4)
#removing the kids in English early under Age 3.4 (youngest age for ASL early) to age match the samples a bit better only removes 3 kids (now 101 instead of 104)
View(GiveN_Early)
```


### Demographics & Basic plots
```{r}
table(GiveN_Early$Language_Modality, GiveN_Early$GN_smto5_fac, useNA = "always") # Table of knower-levels by Language Modality
prop.table(table(GiveN_Early$Language_Modality, GiveN_Early$GN_smto5_fac, useNA = "always"), margin=1) #proportions calculated by row (within language group)

#TO GET AGES/SES etc by group
ggplot(data=GiveN_Early, aes(x=Language_Modality, y=Age)) + geom_violin() + labs(x="Language Modality", y="Age (years)")

ggplot(data=subset(GiveN_Early, (Language_Modality=="ASL")), aes(x=Age, y=..count..)) + geom_density()
ggplot(data=subset(GiveN_Early, (Language_Modality=="English")), aes(x=Age, y=..count..)) + geom_density()

ggplot(data=GiveN_Early, aes(x=Language_Modality, y=SES..8.66.)) + geom_violin() + labs(x="Language Modality", y="SES")

ggplot(data=subset(GiveN_Early, (Language_Modality=="ASL")), aes(x=SES..8.66., y=..count..)) + geom_density()
ggplot(data=subset(GiveN_Early, (Language_Modality=="English")), aes(x=SES..8.66., y=..count..)) + geom_density()


# Give-N by Age/Modality w/tiny jitter
jitter <- position_jitter(width=0.2, height=0.1)
ggplot(data=subset(GiveN_Early, !is.na(LangGrp)), aes(x=Age, y=GNsm_to_5)) + geom_point(aes(shape=Modality, color=Modality), position=jitter, size=2.5) + geom_smooth(mapping=aes(group=Modality, color=Modality), method="loess", se=FALSE) + labs(x="Age (years)", y="Give-N highest number correctly given (up to 5)") + theme(text = element_text(size=11), legend.position = c(0.87, 0.14)) + scale_x_continuous(breaks = seq(3,7, by = 1), limits=c(3,7)) + scale_y_continuous(breaks = seq(0, 5, by = 1), limits=c(0,5))

# Give-N by Age/Modality actual jitter plot
ggplot(data=subset(GiveN_Early, !is.na(LangGrp)), aes(x=Age, y=GNsm_to_5)) + geom_jitter(aes(shape=Modality, color=Modality), size=2.5) + geom_smooth(mapping=aes(group=Modality, color=Modality), method="loess", se=FALSE) + labs(x="Age (years)", y="Give-N highest number correctly given (up to 5)") + theme(text = element_text(size=11), legend.position = c(0.87, 0.14))  + scale_x_continuous(breaks = seq(3,7, by = 1), limits=c(3,7)) + scale_y_continuous(breaks = seq(0, 5, by = 1), limits=c(0,5))
```

### Subsetting the data by age
We have quite a few participants at ceiling on Give-N, likely because of our age range, and this may be obscuring any potential group differences--so we subset to a smaller age range, using information on the distribution of ages
```{r}
##TABLE OF AGES
#set up age bins first
# set up cut-off values 
breaks <- c(3.4, 4, 4.5, 5, 5.5, 6, 6.5, 7, 7.6)
# specify interval/bin labels
tags <- c("[3.4-4)","[4-4.5)", "[4.5-5)", "[5-5.5)", "[5.5-6)", "[6-6.5)","[6.5-7)", "[7-7.6)")
# bucketing values into bins
GiveN_Early$AgeBins <- cut(GiveN_Early$Age, 
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)
# inspect bins
summary(GiveN_Early$AgeBins)

##make the tables
table(GiveN_Early$AgeBins, GiveN_Early$Language_Modality)
prop.table(table(GiveN_Early$AgeBins, GiveN_Early$Language_Modality))

#set up DIFFERENT age bins first
# set up cut-off values 
breaks_fewer<- c(3.4, 4.5, 5.5, 6.5, 7.6)
# specify interval/bin labels
tags_fewer <- c("[3.4-4.5)","[4.5-5.5)", "[5.5-6.5)", "[6.5-7.6)")
# bucketing values into bins
GiveN_Early$AgeBinsFewer <- cut(GiveN_Early$Age, 
                           breaks=breaks_fewer, 
                           include.lowest=TRUE, 
                           right=FALSE, 
                           labels=tags_fewer)
# inspect bins
summary(GiveN_Early$AgeBinsFewer)

##make the tables
table(GiveN_Early$AgeBinsFewer, GiveN_Early$Language_Modality)
prop.table(table(GiveN_Early$AgeBinsFewer, GiveN_Early$Language_Modality))
```

Based on charts and participant, we decided to subset the data to participants under 5.5 years of age
```{r}
GiveN_3_to_5_half <- subset(GiveN_Early, Age <5.6) 
View(GiveN_3_to_5_half)

##Contingency table for do kids know 5 or not by group
GiveN_3_to_5_half <- GiveN_3_to_5_half %>% dplyr::mutate(At_5 = ifelse(GN_Small_Ceiling._FinalSet_conservative>=5, "Yes", "No"))
View(GiveN_3_to_5_half)

##Table of how many kids "are "know 5" (or higher) by group (Contingency table)
table1::label(GiveN_3_to_5_half$At_5) <- "Does child know up to 5?"
table1::table1(~At_5 | Language_Modality, data = GiveN_3_to_5_half)

```
 ***

### QUESTION: Within this age group, do we see different proportions of kids who are 5-knowers vs. not?
```{r}

#Using risk ratio code to see if English group sig different from ASL group
Early_ASL_v_English_know5 <- matrix(c(8,12,19,36), 2, 2, byrow=TRUE)
dimnames(Early_ASL_v_English_know5) <- list("Group" = c("Early ASL", "Early English (Typically Hearing)"), "Does child know up to 5?" = c("No", "Yes"))
Early_ASL_v_English_know5
riskratio.small(Early_ASL_v_English_know5)

```
**Findings**: Risk ratio analysis indicates that Early English participants aged 3.5-5.5 years are equally likely to be 5-knowers as Early ASL participants
<br>

### Alternative analysis: Ordinal logistic regression
Do children learning ASL progress through "knower-levels" at earlier ages/more rapidly than children learning spoken English
```{r}
## Recode the Give-N small variable so 
GiveN_3_to_5_half$GN_upto_5 <- GiveN_3_to_5_half$GN_sm_fac_c
str(GiveN_3_to_5_half$GN_upto_5)
GiveN_3_to_5_half$GN_upto_5 <- dplyr::recode(GiveN_3_to_5_half$GN_upto_5, "6" = "5")
str(GiveN_3_to_5_half$GN_upto_5)
View(GiveN_3_to_5_half)
GN_to_5 <-GiveN_3_to_5_half$GN_upto_5
summary(GN_to_5)

dplyr::filter(GiveN_3_to_5_half, LanguageGroup == "English Early") %>% count() #48 English Early kids in the above graph
dplyr::filter(GiveN_3_to_5_half, LanguageGroup == "ASL Early") %>% count() #27 ASL Early kids in the above graph

#Graph of Give-N scores (up to 6) for English Early participants only
dplyr::filter(GiveN_3_to_5_half, GiveN_3_to_5_half$Group_4cat == "English Early") %>% ggplot(GiveN_3_to_5_half, mapping=aes(x=Age, y=GN_Small_Ceiling._FinalSet_conservative)) + geom_point(color = "grey39") + geom_smooth(method = "lm") + labs(x="Age (years)", y="Give-N Score") + theme(text = element_text(size=16)) + coord_cartesian(xlim = c(3,5.5), ylim= c(0,6))



# Scatterplot of Give-N by age separated by Modality (is slope steeper for ASL than for spoken English?)
ggplot(GiveN_3_to_5_half, aes(x=Age, y=GN_Small_Ceiling._FinalSet_conservative)) + geom_point(aes(shape=Language_Modality, color=Language_Modality), position=jitter, size=2.5) + geom_smooth(mapping=aes(group=Language_Modality, color=Language_Modality), method="lm", se=FALSE) + labs(x="Child's Age (years)", y="Give-N Score") + scale_shape_manual(name="Language", labels=c("ASL (n=27)", "English (n=48)"), values=c(17, 16)) + scale_color_manual(name="Language", labels=c("ASL (n=27)", "English (n=48)"), values=c("mediumvioletred", "grey55")) + theme(text = element_text(size=14), legend.position = c(0.8, 0.2), legend.title = element_text(size=12, face="bold")) + coord_cartesian(xlim = c(3,5.5), ylim= c(0,6))




j <- polr(formula = GN_upto_5 ~ SES..8.66. + Age * Language_Modality, data = GiveN_3_to_5_half, Hess = TRUE)
summary(j)
j$AIC <- AIC(j) #needed to get AIC in output table
stargazer(j, title = "'Iconic' Number Symbols Do NOT Facilitate Number Knowledge for Early-Exposed DHH Children", align=TRUE, dep.var.labels=c("Give-N Highest Quantity Correct (up to 5)"), covariate.labels=c("Socioeconomic Status (SES)", "Age (Years)", "Language Modality (Spoken English)", "Age x Modality Interaction"), single.row=TRUE, keep.stat=c("n", "aic"), out= "GN_to5_SES_Modality*Age__EARLYONLY_210219.htm")

#Question: DO WE HAVE THE POWER TO DETECT a MODALITY EFFECT if one existed?
 # TO DO: Try looking at the confidence intervals of the coefficients


```

